### 2019-07-11 notebook/ghost-fit.Rmd
and so it begins... :camping:

---
title: "greta.Rmd"
author: "Carl Boettiger"
date: "5/30/2019"
output: github_document
---

```{r message = FALSE, warning=FALSE}
library(tidyverse)
library(greta) # remotes::install_github("greta-dev/greta")
set.seed(123456)
```

```{r}
data <- read_csv("../data/reps.csv")
data %>% ggplot(aes(t,x, group=reps)) + geom_line(alpha=0.1)
```

```{r}
wide <- data %>% spread(reps, x) %>% select(-t) %>% 
  select(2) %>% # which / how many reps we use
  as.matrix()
n <- dim(wide)[1]

```

```{r}
x_t1 <- wide[-1,] # data$x[-1]
x_t <- wide[-n,] # data$x[-100]
dim(x_t)
dim(x_t1)
#tibble(x_t, x_t1) %>% ggplot(aes(x_t, x_t1)) + geom_point()
```

```{r}
#r <- 0.05 
Q <- 5
#sigma <- 0.02
#a <- 0.0233
K <- uniform(0, 10)
H <- uniform(0,  2)
a <- uniform(0, 1)
r <- uniform(0, 1)
#Q <- uniform(0, 10)
sigma <- uniform(0, 1)

# Model
mean <- x_t + r * x_t * (1 - x_t / K) - a * x_t ^ Q / (x_t ^ Q + H ^ Q)
distribution(x_t1) <- normal(mean, sigma)
m <- model(K, a, H, sigma, r)
```

```{r }
system.time({
  draws <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
summary(draws)
```

```{r}
#bayesplot::mcmc_trace(draws)
```

```{r}
samples <-  
  map_dfr(draws, 
          function(x) data.frame(x, t = 1:dim(x)[1]), 
          .id = "chain") %>% 
  gather(variable, value, -t, -chain)

samples %>%  
  ggplot(aes(t,value, col=chain, group=chain)) + 
  geom_line() +
  facet_wrap(~variable, scales = "free") + 
  scale_color_viridis_d()
```

```{r}
#Q = 5
true <- data.frame(a = 0.023, r = .05, K = 2, H = .38, sigma = .02) %>%
  gather(variable, value)
```

```{r}
samples %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```