### 2019-07-11 notebook/ghost-sims.Rmd

---
title: "simulation & inference of transients"
author: "Carl Boettiger"
date: "5/30/2019"
output: github_document
---

```{r}
library(tidyverse)
library(nimble)
```

```{r}
#a = 0.023
r = 0.05
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = .02, a=0.023, N = 1e3)
growth <- function(x, p) x * p$r * (1 - x / p$K)
consumption <- function(x,p) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)
```

```{r}
theory <- 
  tibble(x= seq(0,2, length.out = 100)) %>%
  mutate(g = growth(x,p), 
         c = consumption(x,p)) %>%
  mutate(potential = - cumsum(g - c)) %>%
  gather(curve, y, -x, -potential) 
```

```{r noisy_switch}
theory %>%
  ggplot(aes(x, y, col=curve)) +
  geom_line(lwd=1)
```

```{r}
theory %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd=1)
```

```{r}
library(nimble)
library(tidyverse)
# Define stochastic model in BUGS notation
may  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    # Determinstic mean looks like standard R
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)  - a * x[t] ^ Q / (x[t] ^ Q + H ^ Q)
    # Note the use of ~ in BUGS to show 'distributed as normal' 
    y[t+1] ~ dnorm(mu[t], sd = sigma)
    x[t+1] <- max(y[t+1],0)
  }

})
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.2))
cmodel <- model #compileNimble(model)
set.seed(123456)
simulate(cmodel)
df <- tibble(t = seq_along(cmodel$x), x = cmodel$x)
```

```{r}
df %>% ggplot(aes(t,x)) + geom_point()
df %>% write_csv("../data/ex.csv")
```

Let's do 100 replicates

```{r}
set.seed(123)

df <- map_dfr(1:100, 
  function(i){
    simulate(cmodel)
    tibble(t = seq_along(cmodel$x), x = cmodel$x, reps = i)
   })
```

```{r}
df %>% 
  #filter(t %in% seq(1000, 2000, by = 1)) %>% 
  ggplot(aes(t,x, group=reps)) + geom_line(alpha=.2)

df %>% write_csv("../data/reps.csv")
```