### 2019-07-12 notebook/ghost-fit.Rmd
add may outbreak model

---
title: "greta.Rmd"
author: "Carl Boettiger"
date: "5/30/2019"
output: github_document
---

```{r message = FALSE, warning=FALSE}
library(tidyverse)
library(greta) # remotes::install_github("greta-dev/greta")
set.seed(123456)
```

```{r}
data <- read_csv("../data/reps.csv") # %>% filter(t %in% seq(0,4000, by=4))
data %>% ggplot(aes(t,x, group=reps)) + geom_line(alpha=0.02)
```

Divide this data into several groups based on

```{r}
means <- data %>% group_by(reps) %>% summarise(ave = mean(x)) %>% arrange(ave)

lows <- means %>% dplyr::slice(1:10) %>% pull(reps)
highs <- means %>% dplyr::slice((n()-10):n()) %>% pull(reps)
coding <- function(reps) case_when(
  reps %in% lows ~ "low",
  reps %in% highs ~ "high"
)
df <- data %>% mutate(coding = coding(reps)) %>% na.omit()
df %>% ggplot(aes(t, x, group=reps, color = coding)) + geom_line(alpha=0.5)

#means %>% ggplot(aes(ave)) + geom_histogram(binwidth = 0.02)

```

## highs

```{r}
wide <- data %>% 
  filter(reps %in% highs) %>%
  spread(reps, x) %>%
  select(-t) %>% 
  as.matrix()
n <- dim(wide)[1]

```

```{r}
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
```

```{r}
r <- 0.05 
Q <- 5
K <- 2
# a <- 0.0233
H <- 0.38
sigma <- 0.02

# r <- uniform(0, 1)
# Q <- uniform(0, 10)
#K <- uniform(0, 10)
a <- uniform(0, 1)
# H <- uniform(0,  2)
#sigma <- uniform(0, 1)

# Model
mean <- x_t + r * x_t * (1 - x_t / K) - a * x_t ^ Q / (x_t ^ Q + H ^ Q)
distribution(x_t1) <- normal(mean, sigma)
m <- model(a)
```

```{r }
system.time({
  draws <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
summary(draws)
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
samples <-  
  map_dfr(draws, 
          function(x) data.frame(x, t = 1:dim(x)[1]), 
          .id = "chain") %>% 
  gather(variable, value, -t, -chain)
```

```{r include=FALSE}
## manual version of bayesplot::mcmc_trace()
samples %>%  
  ggplot(aes(t,value, col=chain, group=chain)) + 
  geom_line() +
  facet_wrap(~variable, scales = "free") + 
  scale_color_viridis_d()
```

```{r}
#Q = 5
true <- 
  data.frame(a = 0.023 #, 
             #r = .05, 
             #K = 2, 
             #H = .38, 
             #sigma = .02
             ) %>%
  gather(variable, value)
```

```{r}
samples %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

## lows

```{r}
wide <- data %>% 
  filter(reps %in% lows) %>%
  spread(reps, x) %>%
  select(-t) %>% 
  as.matrix()
n <- dim(wide)[1]

```

```{r}
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
```

```{r}
mean <- x_t + r * x_t * (1 - x_t / K) - a * x_t ^ Q / (x_t ^ Q + H ^ Q)
distribution(x_t1) <- normal(mean, sigma * x_t)
m <- model(a)
```

```{r }
system.time({
  draws <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
samples %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```