### 2019-07-19 notebook/may-outbreak-ghost.Rmd
Examples of learning under different priors.
With a wide prior, the decision-maker acts optimally to squash the outbreak early on. The decision-maker never really learns for sure if this was the right decision (i.e. if this was really necessary or if the outbreak would have stopped at low level), it is basically a precautionary reaction -- the posterior belief remains broad since learning opportunity is lost by crushing the outbreak.

In contrast, the decision-maker with a prior more confident in a low-level stable state (i.e. mistaking past evidence of the "ghost" for a stable state), fails t

---
output: github_document
---

```{r}
knitr::opts_chunk$set(fig.width = 7)
```

```{r message=FALSE}
library(MDPtoolbox)
library(sarsop) # remotes::install_github("boettiger-lab/sarsop")
library(tidyverse) # for plotting
library(mdplearning)

#if(interactive()) ggplot2::theme_set(ggdark::dark_theme_light())
#ggplot2::theme_set(theme_grey())
#ggdark::invert_geom_defaults()
```

# Outbreak model

- 'harvest' term corresponds to removal of pest, with associated cost
- also experience damage costs proportional to pest abundance

```{r}
damage <- 0.05
control <- 1
reward_fn <- function(x,h) - damage * x ^ 2 - control * h
discount <- 0.98

states <- seq(0,2, length=200)
actions <- states
observations <- states
sigma_g <- 0.05
sigma_m <- 0.0

r <- 0.8
K <- 1.53
q <- 2
b <- .2
eps <- states[2]/10
Tmax <- 100

may <- function(a){  
  function(x, h){ # May
    # x <- pmax(x - h, 0)   # harvest then recruit   
    x + x * r * (1 - x / K)  - a * x ^ q / (x ^ q + b ^ q) + eps - h
  }
}

```

Range of possible a that covers tipping in both directions:

```{r}
possible_a <- seq(.25, .34, by = 0.005)
true_a <- 0.27   ## reality has just a transient.  use  0.277 for stronger ghost
believe_a <- 0.29 ## believe there's an attractor
true_i <- which.min(abs(possible_a - true_a))
```

```{r}
f <- may(true_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
f <- may(believe_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
m_true <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(true_a), sigma_g, sigma_m, noise = "lognormal")

m <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(believe_a), sigma_g, sigma_m, noise = "lognormal")
```

```{r, results="hide"}
soln <- mdp_value_iteration(m$transition, m$reward, discount)
opt_soln <- mdp_value_iteration(m_true$transition, m_true$reward, discount)

```

Policy based on the belief (i.e. that system is bi-stable)

```{r}
df <- tibble(state = states,
             action = actions[soln$policy],
             value = soln$V)
df %>% ggplot(aes(state, action)) + geom_point() 
```

Optimal policy (knowing it is really just a transient)
```{r}
tibble(state = states,
       action = actions[opt_soln$policy]) %>% 
  ggplot(aes(state, action)) + geom_point() 
```

```{r}
x0 <- which.min(abs(states - 0.1))
```

```{r}
no_policy <- numeric(length(states)) + 1

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Result experienced by incorrect belief:  initial in-action followed by need for continued maintenance to prevent high-level outbreak:

```{r}

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Expected result based on the belief: stable low level is acceptable, so no action is required:

```{r}

df <- mdp_planning(m$transition, m$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Optimal strategy knowing this is just a transient (will depend on discount rate):

```{r}

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = opt_soln$policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

```{r}

models <- map(possible_a, function(a){
  fisheries_matrices(states, actions, observations, reward_fn, 
                     may(a), sigma_g, sigma_m, noise = "lognormal")
})

```

```{r}
transition <- lapply(models, `[[`, "transition")
reward <- models[[1]][["reward"]]
```

## a near ghost

```{r}
prior <- dnorm(possible_a, believe_a, 0.01)
prior <- prior / sum(prior)
```

```{r}
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity") +
  geom_vline(aes(xintercept = true_a), col="red", lwd=1, lty=2) + 
  geom_vline(aes(xintercept = believe_a), col="blue", lwd=1, lty=2) 
```

```{r}
sim <- mdp_learning(transition, reward, discount, 
                    x0 = x0, 
                    Tmax = Tmax, 
                    true_transition = transition[[true_i]],
                    model_prior = prior,
                    type = "value iteration", 
                    epsilon = 1e-2)
```

```{r}
 sim$df %>% 
  select(-value) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, states[state], color = series)) + geom_line()
```

```{r}
 sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_line()
```

```{r}
```