### 2019-07-19 slides/boettiger.Rmd
WIP

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "custom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(img/royalsociety/penguin-swim.jpg)
background-position: center
background-size: 120%
class: center, top

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Decisions vs transient dynamics: what could possibly go wrong?

???

Image credit: [Royal Society Photo Competition](https://royalsociety.org/journals/publishing-activities/photo-competition/)

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .6), rgba(150, 150, 150, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, middle, inverse

# in an era of predictive black-box models...

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .6), rgba(150, 150, 150, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, middle, inverse

# ... a need for theory & process-based models

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-size: 120%
class: center, middle, inverse

# Ghost Attractors

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-size: 120%
class: center, top, inverse

# Ghost Attractors

![](img/ghost-invert.png)

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/reef-shark.jpg)
background-size: 120%
class: center, top, inverse

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
class: center, top, inverse

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, top, inverse

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/gentoo-penguin.jpg)
background-size: 120%
class: center, top, inverse

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/eagle-ray.jpg)
background-size: 120%
class: center, top, inverse

---
class: center, middle
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/bohemian-waxwing.jpg)
background-size: 120%

# Acknowledgements

![](img/hastings.jpg)

![](img/nimbios.jpg)

<!--
- Lizzie Wolkovich
- Henry Scharf
-->