### 2019-08-06 notebook/discrete-sims.Rmd
nice simple example of discrete simulation

---
output: github_document
---

```{r}
knitr::opts_chunk$set(fig.width = 7)
```

```{r message=FALSE}
library(MDPtoolbox)
library(mdplearning)
library(expm)
library(truncnorm)
library(tidyverse)

```

# Outbreak model

- 'harvest' term corresponds to removal of pest, with associated cost
- also experience damage costs proportional to pest abundance

```{r}
damage <- 0.05
control <- 1
reward_fn <- function(x,h) - damage * x ^ 2 - control * h
discount <- 0.98

n_s <- 200
states <- seq(0,2, length=n_s)
actions <- states
observations <- states
sigma_g <- 0.05

r <- 0.8
K <- 1.53
q <- 2
b <- .2
eps <- states[2]/10
Tmax <- 100

may <- function(a){  
  function(x, h = 0){ # May
    # x <- pmax(x - h, 0)   # harvest then recruit   
    x + x * r * (1 - x / K)  - a * x ^ q / (x ^ q + b ^ q) + eps - h
  }
}

```

Range of possible a that covers tipping in both directions:

```{r}
possible_a <- seq(.25, .34, by = 0.005)
true_a <- 0.27   ## reality has just a transient.  use  0.277 for stronger ghost
## 0.28: technically 'ghost' is an attractor but effectively not
believe_a <- 0.28 ## believe there's an attractor
true_i <- which.min(abs(possible_a - true_a))
```

```{r}
f <- may(true_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
f <- may(believe_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
transition_matrix <- function(states, f, sigma){
    n_s <- length(states)
    transition <- array(0, dim = c(n_s, n_s))
    for (k in 1:n_s) {

            nextpop <- f(states[k])
            if (nextpop <= 0) {
                transition[, k] <- c(1, rep(0, n_s - 1))
            }
            else {
              x <- truncnorm::dtruncnorm(states, 0, max(states), nextpop, sigma * nextpop)
              transition[, k] <- x / sum(x)

            }

    }
    transition
}
```

```{r}
m_true <- transition_matrix(states, may(true_a), sigma_g)
m <- transition_matrix(states, may(believe_a), sigma_g)
```

"simulate" without any actions

```{r}
x0 <- which.min(abs(states - 0.1))
Tmax <- 300
```

```{r}
X <- numeric(length(states))
X[x0] <- 1

prob_dynamics <- function(M, X, Tmax){
  probability <- M %^% Tmax %*% X 
  data.frame(state = states, probability)
}

prob_dynamics(m, X, Tmax) %>% 
  ggplot(aes(state,probability))  + geom_point() # + geom_bar(stat="identity")
```

```{r}

sim <- function (transition,  x0, Tmax ){
    n_states <- dim(transition)[2]
    state <- numeric(Tmax + 1)
    state[1] <- x0
    time <- 1:Tmax
    for (t in time) {
        state[t + 1] <- base::sample(1:n_states, 
                               1, 
                               prob = transition[, state[t]])
    }
    data.frame(time = 1:Tmax, state = state[time])
}

sim(m, x0, Tmax) %>% 
  mutate(state = states[state]) %>% 
  ggplot(aes(time, state)) + geom_point() + geom_path() 
```

```{r}
```