### 2019-09-23 manuscript/appendix/appendix.Rmd
WIP
Need to revise & consolidate notebooks

---
title: "Appendix"
author: "Carl Boettiger"
output: github_document
---

```{r message = FALSE}
library(tidyverse)
library(patchwork)
library(furrr)
set.seed(12345)
```

## Model Definition

```{r}
p <- list(r = .05, K = 200, Q = 5, H = 38, sigma = .02, a=2.3, N = 3e3, x0 = 20)
```

```{r}
growth <- function(x, p) x * p$r * (1 - x / p$K)
consumption <- function(x, p) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)
may <- function(x, p) x + growth(x,p) - consumption(x,p)
```

```{r}
theory <- 
  function(p){
    tibble(x= seq(0,p$K, length.out = 100)) %>%
    mutate(growth = growth(x,p), 
           consumption = consumption(x,p)) %>%
    mutate(potential = - cumsum(growth - consumption)) %>%
    gather(curve, y, -x, -potential) 
}
```

Plots of the growth rate functions and the potential well diagram

```{r}
fig_a <- theory(p) %>%
  ggplot(aes(x, y, col=curve)) +
  geom_line(lwd=1) + 
  labs(title = "Population dynamics") +
  theme(legend.position = c(.5, .2), legend.background = element_blank())

fig_b <- theory(p) %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd=1)  + 
  labs(title = "Potential Well Diagram")

fig_a + fig_b
```

## Model simulations

```{r}
# Generic simulator routine
sim <- function(f, p){
  x <- numeric(p$N)
  x[1] <- p$x0
  dBt <- numeric(p$N)
  if(p$sigma > 0) dBt <- rnorm(p$N, 0, p$sigma)
  for(t in 1:(p$N-1)){
    x[t+1] <- max(f(x[t], p) + x[t] * dBt[t], 0)
  }
  data.frame(t = 1:p$N, x)
}
```

Deterministic simulations:

```{r}
p_det <- p
p_det$sigma <- 0
det <- sim(may,p_det) 
det %>% ggplot(aes(t, x)) + geom_line() 
```

Compare deterministic simulations across a range of `a` values:

```{r}
a <- c(2.25, 2.30, 2.305, 2.31)
names(a) <- a
p_det$N <- 1e4
df <- map_dfr(a, function(a){
  p_det$a <- a
  sim(may, p_det)
},
.id = "a")

df %>% ggplot(aes(t,x, col = a)) + geom_line() + coord_cartesian(xlim=c(0,5000))

```

Stochastic replicate simulations (run in parallel):

```{r}
plan("multisession")
system.time({
  df <- future_map_dfr(1:500, function(reps) sim(may, p), .id = "reps")
})

```

```{r}
p_det$N <- 3e3
det <- sim(may,p_det) %>% mutate(reps=1)
mean <- df %>% group_by(t) %>% summarise(x = mean(x)) %>% mutate(reps = 1)
```

## Stochastic ensemble mean vs deterministic mean

```{r}
df %>% group_by(reps) %>%
  sample_n(5e2) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(alpha=0.05) +
  geom_line(data = det, color = "blue") + 
  geom_line(data = mean, color = "red")
```

---

# ARIMA Forecast

```{r}
library(tidyverse)
library(stats)
```

Extract a series with relatively long ghost:

```{r}
means <- df %>% group_by(reps) %>% summarise(ave = mean(x), max = max(x))
lows <- means %>% filter(ave < mean(ave) - 3*sd(ave), max > 120) %>% pull(reps)
```

```{r arima_forecast}
x <- df %>% filter(reps == lows[[1]], t <= 2000) %>% pull(x) %>% as.ts() 
fit <- arima(x)
y <- predict(fit, n.ahead = 1000) %>% as.data.frame()
y <- bind_rows(data.frame(pred = x, se = 0), y)

arima_forecast <- df %>% 
  filter(reps == lows[[1]]) %>% 
  select(t,x) %>%
  bind_cols(y)
```

```{r arima_figure, message=FALSE, warning=FALSE}
train <- arima_forecast %>% filter(t < 2000) 
test <- arima_forecast %>% filter(t > 2000) 
train %>%
  ggplot(aes(t)) + 
  geom_line(aes(y=x)) +
  geom_line(aes(y=pred), lwd = 1, lty = 3, col = "red", data = test) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se), alpha = 0.4, fill = "red", data = test)  + 
  geom_line(aes(y=x), data = test, lty = 2, col="blue") +
  xlim(0,3000) + ylab("x") + theme_minimal()

```

----

# Model-based forecast

```{r}
library(greta)
```

Bayesian estimate on training data only:

```{r}
wide <- select(train, x) %>% as.matrix() 
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
```

```{r}
r <- p$r
Q <- p$Q
K <- p$K
# a <- 2.3
H <- p$H
sigma <- p$sigma

# r <- uniform(0, 1)
# Q <- uniform(0, 10)
#K <- uniform(0, 10)
a <- uniform(0, 10)
# H <- uniform(0,  2)
#sigma <- uniform(0, 1)

# Model   (mean <-  may(x_t, p))
mean <- x_t + r * x_t * (1 - x_t / K) - a * x_t ^ Q / (x_t ^ Q + H ^ Q)
distribution(x_t1) <- normal(mean, sigma * x_t)
m <- model(a)
```

```{r }
system.time({
  draws <- mcmc(m, n_samples = 1000, warmup = 3000, chains = 4, verbose = FALSE)
})
```

```{r}
samples <-  
  map_dfr(draws, 
          function(x) data.frame(x, t = 1:dim(x)[1]), 
          .id = "chain") %>% 
  gather(variable, value, -t, -chain)
```

```{r}
#Q = 5
true <- 
  data.frame(a = 2.3) %>%
  gather(variable, value)
```

```{r}
samples %>% ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")

```

Replicate simulations of stochastic model with parameters drawn from posteriors

(would be great to have a `greta` method for this...)

```{r}
a <- sample(unlist(draws, TRUE), 10)
p$N <- 3e3
posterior_sims <- map_dfr(a, function(a){
  p$a <- a
  sim(may, p)
},
.id = "a")

```

```{r}
posterior_sims %>% 
  mutate(a = as.numeric(a)) %>% 
  ggplot(aes(t,x, group=a)) + 
  geom_line(alpha = 0.05) 
```

```{r}
```

----