### 2019-10-25 notebook/gp-manual-vs-greta.Rmd
WIP

---
layout: post
title: Basic regression in Gaussian processes 
category: ecology
tag: nonparametric-bayes
code: true
output: github_document
---

```{r}
library(tidyverse)
library(greta)
library(greta.gp)

require(MASS)
require(reshape2)
require(ggplot2)
set.seed(12345)
```

```{r}
x_predict <- seq(-5,5,len=50)
l <- 1
```

We will use the squared exponential (also called radial basis or Gaussian, though it is not this that gives Gaussian process it's name; any covariance function would do) as the covariance function,

$$\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$$

```{r}
SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
cov <- function(X, Y) outer(X, Y, SE, l)
```

```{r}
COV <- cov(x_predict, x_predict)
```

Generate (draw) a number of functions from the process

```{r}
values <- mvrnorm(3, rep(0, length=length(x_predict)), COV)
```

Reshape the data into long (tidy) form, listing x value, y value, and sample number

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

fig2a <- ggplot(dat,aes(x=x,y=value)) +
  geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
  geom_line(aes(group=variable))  +
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x") +
  theme_bw()
fig2a
```

### Posterior distribution given the data

Imagine we have some data,

```{r}
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
```

In general we aren't interested in drawing from the prior, but want to include information from the data as well.  We want the joint distribution of the observed values and the prior is:

$$\begin{pmatrix} y_{\textrm{obs}} \\ y_{\textrm{pred}} \end{pmatrix} \sim \mathcal{N}\left( \mathbf{0}, \begin{bmatrix} cov(X_o,X_o) & cov(X_o, X_p) \\ cov(X_p,X_o) & cov(X_p, X_p) \end{bmatrix} \right)$$

## No observation noise

Assuming the data are known without error and conditioning on the data, and given $x \sim \mathcal{N}(0, cov(X_o, X_o))$, then the conditional probability of observing our data is easily solved by exploiting the nice properties of Gaussians,

$$x|y \sim \mathcal{N}\left(E,C\right)$$
$$E = cov(X_p, X_o) cov(X_o,X_o)^{-1} y$$
$$C= cov(X_p, X_p) - cov(X_p, X_o) cov(X_o,X_o)^{-1} cov(X_o, X_p) $$

(We use `solve(M)` which with no second argument will simply invert the matrix `M`, but should use the Cholsky decomposition instead for better numerical stability)

```{r}
cov_xx_inv <- solve(cov(obs$x, obs$x))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

Now lets take 3 random samples from the posterior distribution,

```{r}
values <- mvrnorm(3, Ef, Cf)
```

and plot our solution (mean, 2 standard deviations, and the ranom samples.)

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

gp <- data.frame(x = x_predict, Ef = Ef, sigma = 2*sqrt(diag(Cf)) )

ggplot(dat,aes(x=x,y=value)) + 
  geom_ribbon(data = gp, 
              aes(x, 
                  y = Ef, 
                  ymin = Ef - sigma, 
                  ymax = Ef + sigma
              ),
              fill="grey80") +
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(dat = gp, aes(x=x,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
```

![plot of draws from the posterior, with no proces noise](http://farm9.staticflickr.com/8237/8591045344_918138a53a_o.png)

Additive noise
--------------

In general the model may have process error, and rather than observe the deterministic map $f(x)$ we only observe $y = f(x) + \varepsilon$.  Let us assume for the moment that $\varepsilon$ are independent, normally distributed random variables with variance $\sigma_n^2$.

```{r}
sigma.n <- 0.25
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

Now lets take 3 random samples from the posterior distribution,

```{r}
values <- mvrnorm(3, Ef, Cf)
```

and plot

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

gp <- data.frame(x_predict = x_predict, Ef = Ef, sigma = 2*sqrt(diag(Cf)) )

fig2c <- ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=gp, 
              aes(x=x_predict, y=Ef, ymin=Ef-sigma, ymax=Ef+sigma),
              fill="grey80") + # Var
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(data=gp,aes(x=x_predict,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2c
```

Note that unlike the previous case, the posterior no longer collapses completely around the neighborhood of the test points.