### 2019-10-25 notebook/greta-gp.Rmd

---
output: github_document

---

```{r setup}
library(tidyverse)
library(greta)
library(greta.gp)
#tensorflow::use_session_with_seed(1234)

```

```{r}
arima_forecast <- readr::read_csv("../data/arima_forecast.csv.gz")
train <- arima_forecast  %>% filter(t %in% seq(0, 2000, by = 5))

wide <- select(train, x) %>% as.matrix() 
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,]

y <- c(0,x_t1 - x_t)
x <- c(0, x_t)
x_plot <- seq(0, 2, length.out = 200)
```

```{r}

```

```{r}
# simulate toy data
#x <- runif(20, 0, 10)
#y <- sin(x) + rnorm(20, 0, 0.5)
#x_plot <- seq(-1, 11, length.out = 200)
```

```{r}

# hyperparameters
rbf_var <- lognormal(0, 1)
rbf_len <- lognormal(0, .01)
obs_sd <- lognormal(0, .01)

# kernel & GP
kernel <- rbf(rbf_len, rbf_var) + bias(1)
f <- gp(x, kernel)

# likelihood
distribution(y) <- normal(f, obs_sd)

# prediction
f_plot <- project(f, x_plot)
```

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f_plot)
draws <- mcmc(m, one_by_one = TRUE)
```

```{r}
# plot 200 posterior samples
#plot(y ~ x, pch = 16, col = grey(0.4), xlim = c(0, 10), ylim = c(-2.5, 2.5))

plot(y ~ x, pch = 16, col = grey(0.4), xlim = c(0, 2), ylim = c(-0.1, 0.15))
for (i in 1:200) {
  lines(draws[[1]][i, ] ~ x_plot,
        lwd = 2,
        col = rgb(0.7, 0.1, 0.4, 0.1))  
}
```

```{r}
```