### 2019-10-26 notebook/python-reticulate-gp.Rmd

---
output: github_markdown
---

```{python}
import numpy as np
import tensorflow as tf
import tensorflow_probability as tfp
```

```{python}
tfd = tfp.distributions
psd_kernels = tfp.positive_semidefinite_kernels
```

```{python}
num_points = 100
# Index points should be a collection (100, here) of feature vectors. In this
# example, we're using 1-d vectors, so we just need to reshape the output from
# np.linspace, to give a shape of (100, 1).
index_points = np.expand_dims(np.linspace(-1., 1., num_points), -1)

# Define a kernel with default parameters.
kernel = psd_kernels.ExponentiatedQuadratic()
```

```{python}
gp = tfd.GaussianProcess(kernel, index_points)

samples = gp.sample(10)
# ==> 10 independently drawn, joint samples at `index_points`

noisy_gp = tfd.GaussianProcess(
    kernel=kernel,
    index_points=index_points,
    observation_noise_variance=.05)
noisy_samples = noisy_gp.sample(10)

noisy_samples
```