### 2019-10-28 notebook/greta-gp.Rmd

---
output: github_document

---

```{r setup}
library(tidyverse)
library(greta)
library(greta.gp)
#tensorflow::use_session_with_seed(1234)

```

## Model Definition

```{r}
p <- list(r = .05, K = 200, Q = 5, H = 38, sigma = .002, a=2.3, N = 3e3, x0 = 6)
```

```{r}
growth <- function(x, p) x * p$r * (1 - x / p$K)
consumption <- function(x, p) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)
may <- function(x, p) x + growth(x,p) - consumption(x,p)
```

## Model simulations

```{r}
# Generic simulator routine
sim <- function(f, p){
  x <- numeric(p$N)
  x[1] <- p$x0
  dBt <- numeric(p$N)
  if(p$sigma > 0) dBt <- rnorm(p$N, 0, p$sigma)
  for(t in 1:(p$N-1)){
    x[t+1] <- max(f(x[t], p) + x[t] * dBt[t], 0)
  }
  data.frame(t = 1:p$N, x)
}
```

Deterministic simulations:

```{r}
p_det <- p
p_det$sigma <- 0
det <- sim(may,p_det) 
det_diffs <- det %>% mutate(y = lead(x)-x, case = "det")
det_diffs <- det %>% mutate(y = c(diff(x),NA), case = "det")

ex <- sim(may,p) 
ex %>% mutate(y = lead(x)-x, case = "stoch")  %>% bind_rows(det_diffs) %>%
  ggplot(aes(x,y, col=case)) + geom_point()
```

```{r}
ghost <- readr::read_csv("../data/ghost_sims.csv.gz")
wide <- ghost %>% pivot_wider(names_from = reps, values_from = x) %>% select(-t)
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
y <- x_t1 - x_t
Y <- y %>% apply(1, mean, na.rm=TRUE)
X <- x_t %>% apply(1, mean, na.rm=TRUE)

tibble(x=X,y=Y, case="mean") %>% bind_rows(det_diffs) %>% ggplot(aes(x,y, col=case)) + geom_point()

ghost %>% group_by(reps) %>%  mutate(y = lead(x)-x) %>% 
  ggplot(aes(x,y)) + geom_point(size=.1, alpha=0.005)
```

```{r}
arima_forecast <- readr::read_csv("../data/arima_forecast.csv.gz")
train <- arima_forecast  #%>% filter(t %in% seq(0, 2000, by = 5))

wide <- select(train, x) %>% as.matrix() 
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,]

y <- c(0,x_t1 - x_t)
x <- c(0, x_t)
x_predict <- seq(0, 2, length.out = 200)

tibble(x,y) %>% ggplot(aes(x,y)) + geom_point()
```

----

```{r}
# kernel & GP
y <- diff(ex$x)
x <- ex$x[-n]
ggplot(data.frame(x,y), aes(x,y)) + geom_point()

# hyperparameters
rbf_var <- lognormal(0, 1)
rbf_len <- lognormal(0, 10)
obs_sd <- lognormal(0, .1)

# kernel & GP
kernel <- rbf(rbf_len, rbf_var)
f <- gp(x, kernel)

# likelihood -- not clear how this influences f_plot?
distribution(y)<- normal(f, obs_sd)

# prediction
f_plot <- project(f, x_predict)
```

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f_plot, rbf_len, rbf_var, obs_sd)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```

```{r}
df <- draws[[1]] %>% as_tibble(.name_repair = "universal") 
hyperpars <- df %>% dplyr::select(rbf_len, obs_sd, rbf_var)  
posterior <- df %>% dplyr::select(-rbf_len, -obs_sd, -rbf_var)

hyperpars %>% colMeans()
#hyperpars %>% gather() %>% ggplot(aes(x = value, fill=key)) + stat_density() + facet_wrap(~key)
```

```{r}
posterior %>% t() %>% as_tibble() %>% dplyr::select(1:200) %>% mutate(x = x_predict) %>% pivot_longer(-x)  %>%
  ggplot(aes(x,value)) + 
  geom_line(aes(group=name), alpha=0.3, col=rgb(0.7, 0.1, 0.4, 0.1)) +
  geom_point(data = obs, aes(x,y))
```

```{r}
```