### 2019-12-30 notebook/nnetar.Rmd
neural net

---
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(fable)
library(ggthemes)
pal <- solarized_pal()(6)

```

```{r}
n_s <- 121
states <- seq(0, 120, length = n_s)
actions <- seq(0, 120, length = n_s)

## Model constants -- used to compute transistion probabilities
efficiency <- .4    
p <- list(r = .8, K = 153, q = 2, b = 20, sigma = .05, x0 = 20) # fixed parameters

may <- function(a){  
  function(x, h = 0){ # May
    y <- x - efficiency * h
    pmax(
      ## controlling h is controlling the bifurcation point directly...
      y + y * p$r * (1 - y / p$K)  - a * y ^ p$q / (y ^ p$q + p$b ^ p$q),  
      0)
  }
}
```

```{r}
transition_matrix <- function(states, actions, f, sigma){
    n_s <- length(states)
    n_a <- length(actions)
    transition <- array(0, dim = c(n_s, n_s, n_a))
    for(i in 1:n_a){
      for (k in 1:n_s) {
          nextpop <- f(states[k], actions[i])
          if (nextpop <= 0) {
            x  <- c(1, rep(0, n_s - 1))
          } else {
            x <- truncnorm::dtruncnorm(states, 0, max(states), nextpop, sigma * nextpop)
            if(sum(x) <= 0){ 
              x  <- c(1, rep(0, n_s - 1))
            } else {
              x <- x / sum(x)
            }
          }
          transition[k, , i] <- x
      }
  }
  if(any(is.na(transition))) stop("error creating transition matrix")
transition
}
```

```{r}
## `actions` is an argument of indices, not necessarily the action values themselves 
sim <- function (transition,  x0, Tmax, action = rep(1, Tmax)){
    n_states <- dim(transition)[2]
    state <- numeric(Tmax + 1)
    state[1] <- x0
    time <- 1:Tmax
    for (t in time) {
        state[t + 1] <- base::sample(1:n_states, 
                               1, 
                               prob = transition[state[t], , action[t]])
    }
    data.frame(time = 1:Tmax, state = state[time])
}
```

```{r}
P_ghost <- transition_matrix(states, actions, may(27.2), p$sigma)
P_weak_attractor <- transition_matrix(states, actions, may(27.6), p$sigma)
P_bistable <- transition_matrix(states, actions, may(28), p$sigma)
```

```{r}
x0 <- which.min(abs(states - p$x0))
Tmax <- 200
set.seed(12345)
reps <- 50
ghost_sim <- map_dfr(1:reps, function(i)
  sim(P_ghost, x0, Tmax)  %>% mutate(state = states[state], scenario = "ghost"),
  .id = "rep")
weak <- map_dfr(1:reps, function(i)
  sim(P_weak_attractor, x0, Tmax)  %>% mutate(state = states[state], scenario = "bistable"),
  .id = "rep")
bistable  <- map_dfr(1:reps, function(i)
  sim(P_bistable, x0, Tmax)  %>% mutate(state = states[state], scenario = "bistable"),
  .id = "rep")
```

```{r}
#fig1_sims <- read_csv("../decisions-vs-transients/manuscript/manuscript_cache/fig1_sims.csv")
#bistable <- fig1_sims %>% filter(scenario == "bistable")
rep1 <- weak %>% filter(rep == 2, time <= 100) %>% select(time, state)
```

```{r}
rep1 %>%
  ggplot(aes(time, state)) + 
  geom_line(lwd=1.5, col = pal[1])

```

---

````{r}
horizon <- 200
marble <- rep1 %>%
    as_tsibble(index = time) %>%
  model(nn = NNETAR(state))
fable <- marble %>% 
  forecast(h = horizon, PI=TRUE) 
fc <- fable %>%
  mutate(sd = map_dbl(.distribution, function(x) sd(x[[1]][[1]])))

rep1 %>% filter(time <= 100) %>%
  ggplot(aes(time, state)) + 
  geom_line(lwd=1.5, col = pal[1]) + 
  geom_line(data = fc, aes(time, state), lwd=1.5) + 
  geom_ribbon(data = fc, aes(time, state, ymin = state - 2 *sd, ymax = state + 2 * sd ), 
              col=NA, alpha = 0.2) + 
  coord_cartesian(xlim= c(0,100+ horizon), ylim = c(0,60))

```